Supplementary Report for the article *Functional diversity can facilitate the collapse of an undesirable ecosystem state*

### 2022-11-21 reports/supplement/supplement_0.9.1.Rmd
mostly removing nonstandard stable state data

The rate of change in population densities is given by the growth rate - death rate ([Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x)). Growth rate is the growth rate in absence of any inhibition multiplied by the amount of inhibition.

@@

(ref:repl-meth2) Stable states found with the **replication method**. The left column of panels refers to the stable state when oxygen diffusivity is at first high (favouring the oxic state). The right column refers to when oxygen diffusivity is at first low (favouring the anoxic state). Points (which sometimes are so close as to appear as thick lines) show the stable state at the end of each period of constant oxygen diffusivity. Thin lines are for visualisation only, and join the points. Grey arrows show the oxygen diffusivity at which the system flips, and the direction of the flip. The length of the grey arrows is without meaning.

@@

sim_length <- 1e6 ## currently a problem in how sim length is parsed by the data processing step
num_strains <- 9
plot_here <- all_stab %>%
  filter(simlength == sim_length)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(sim_length,
                                                           " simulation length,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e6
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e6
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e5
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

wait_time <- 1e4
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

```{r div-temp-meth4, echo = FALSE,  include=TRUE, fig.cap='(ref:div-temp-meth4)', fig.height=7, fig.width=6, fig.align="center"}
wait_time <- 1e3
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

(ref:div-temp-meth5) Effects of diversity with wait time of 1e2 hours. (All else as in Figure \@ref(fig:div-repl-meth1)).

```{r div-temp-meth5, echo = FALSE,  include=TRUE, fig.cap='(ref:div-temp-meth5)', fig.height=7, fig.width=6, fig.align="center"}

wait_time <- 1e2
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time)
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

The results below show the stable states as a function of oxygen diffusivity for a system with an intermediate level of diversity in all functional groups, with either 1'000'000 hours at each oxygen diffusivity level (Figure \@ref(fig:ss-temp-meth3)), or 2'000'000 hours (Figure \@ref(fig:ss-temp-meth4)). This shows that 1'000'000 hours duration of the simulations are sufficient to reach stable state, except very close to state transitions on the trajectory of recovery (i.e.: cyanobacteria: trajectory of increasing oxygen diffusivity; sulfate-reducing bacteria: trajectory of decreasing oxygen diffusivity). Very close to these transitions, even at 2'000'000 hours, the dynamics are still transient, with high tolerance strains at high density, even though they will eventually be outcompeted by low tolerance (higher growth rate) strains.

@@

@@

(ref:pbdiv-only-3) Concentrations of three substrates (O: oxygen; SO: oxidised sulfur; SR: reduced sulfur) in a low PB diversity simulation (x-axis) versus substrate concentration in a high PB diversity simulation (y-axis). The figures show only a small range of oxygen diffusivity on the trajectory of increasing oxygen diffusivity. The black line is the 1:1 line.

@@

dd %>%
  filter(State_variable %in% c("Oxygen", "Sulfide", "Oxidised sulfur"),
         direction == "up",
         a_O < as.numeric(aO_flip_sim2),
         a_O > -3.5) %>%
  ggplot(aes(x = log10(Quantity_sim1),
                 y = log10(Quantity_sim2),
                 col = a_O )) +
  geom_point(lwd = 3) +
  geom_path() +
  scale_color_gradient(low = "yellow", high = "red", na.value = NA,
                       breaks = labs,
                       labels = labs) +
  #coord_trans(x = "log10", y = "log10") +
  facet_wrap( ~ State_variable, scales = "free") +
  geom_abline(intercept = 0, slope = 1) +
  xlab(expression(atop(log[10](concentration)~(µM), Low~PB~diversity))) +
  ylab(expression(atop(log[10](concentration)~(µM),High~PB~diversity))) +
  labs(col=expression(paste("log"[10],"(oxygen diffusivity) (h"^{-1}, ")"))) +
  theme_bw() +
  theme(legend.position="top")
```

@@

Owen, please add here 3 figures that match the text above: 
(1) medium variation in only PB showing coexistence of all groups at low ox diff, 
(2) high variation in only PB showing coexistence of all groups at low ox diff and CB never collapsing on the decreasing ox diff trajectory, 
(3) high variation in both CB and PB.

fig_div_vs_o2diff_multistrain(all_stab, which_strains,
                              which_wait_time, figure_title)
```

@@

(ref:samp-effect1) A simple phenomenological model of the effect of number of strains in an ecosystem on the resilience of that ecosystem, when the resilience of a ecosystem is linearly proportional to the maximum trait value of the strains present in that ecosystem. Each data point is a different ecosystem, with the strains in each ecosystem drawn at random from a pool of 20 strains. Each of the 20 strains was assigned a trait value drawn at random from a uniform distribution with minimum 0 and maximum 1. The blue line is the fit of a generalised additive model and the grey ribbon shows the 1 standard error confidence range.

```{r samp-effect1, echo = FALSE,  include=TRUE, fig.cap='(ref:samp-effect1)', fig.height=5, fig.width=6, fig.align="center"}
if(!("tidyverse" %in% .packages()))
  library(tidyverse)
set.seed(12345)
number_of_species <- 20
species_trait <- tibble(species_name = paste0("Species-", letters[1:20]),
                        trait_value = runif(number_of_species))

@@

We investigated if the diversity effects observed in our study were in line with the sampling effect, that is, driven entirely by the presence of the most tolerant strain or if the less tolerant strains played a role as well. To this end, we ran additional simulations where functional groups contained only the strain with greatest tolerance and lowest maximum growth rate. That is, rather than increasing diversity we only increased tolerance. Specifically, for each of the 20 levels of trait variation, we made simulations where each functional group had only one strain, and its trait values corresponded with those of the most tolerant strain for the given level of trait variation. We also manipulated in which of the functional groups tolerance increased. For example, in a simulation to be compared with variation in only the SB group (and no variation in the CB or PB groups), the SB group contained the strain with highest tolerance, whereas CB and PB each contained the strain with mean tolerance. We then compared the tipping point positions resulting from this tolerance manipulation with the tipping point positions resulting from the diversity manipulation presented in the main text. In addition, we compared the strain dynamics across the oxygen diffusivity gradient of the tolerance manipulation and the diversity manipulation. If the observed diversity effects are caused by the sampling effect, we expect there to be no difference between the tolerance manipulation and the diversity manipulation.

To exclude that differences between the tolerance manipulation and the diversity manipulation were driven by the difference in number of strains, we also performed simulations in which we retained nine strains in each functional group but set the tolerance of all nine strains to the maximum tolerance value. Results were the same as those presented below from simulations with one strain.

@@

## remove data where there are no stable states
## this subset is identified by inspection of the figure
#all_stab <- all_stab %>%
#  filter(!(var_treat == "CB-SB" & stand_var > 0.84 & num_strains == -1))

plot_1 <- fig_div_vs_o2diff_maxtolcomp(all_stab, which_strains,
                              which_wait_time, figure_title)
plot_1 +
  xlab("Standardised amount of trait variation used to create nine strains\nof which either all nine are retained, or only the most tolerant is retained.") +
  guides(col=guide_legend(title="Which strains are retained", title.position = "top"))
```

@@

# Comparing the effects of different treatments

(ref:comp-comb) Another method for comparing the effects of different diversity treatments. Some lines (treatment combinations) are over-plotted and therefore show identical effects of some combinations.

```{r comp-comb, echo = FALSE,  include=TRUE, fig.height=8, fig.width=8, fig.align="center", fig.cap='(ref:comp-comb)'}

@@

Increased CB diversity causes a slight decrease in the resilience of the anoxic state. The following two figures show the state responses to oxygen diffusivity with low Figure \@ref(fig:ss-CB-only1) and high Figure \@ref(fig:ss-CB-only2) levels of CB diversity.

The difference in the position of the anoxic to oxic tipping point is small. It is, in fact the smallest it could be, as it is the smallest step size in oxygen diffusivity used in the simulations. Furthermore, we think the presence of higher strain diversity can increase the rate of response of the system (since there is a strain with higher growth rate than in low diversity levels) and that this difference in response rates is likely responsible for the very small difference in resilience.

>>>>

The rate of change in population densities is given by the per capita growth rate - per capita death rate, multipled by population size ([Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x)). Growth rate is the growth rate in absence of any inhibition multiplied by the amount of inhibition.

@@

(ref:repl-meth2) Stable states found with the **replication method**. The left column of panels refers to the stable state when cyanobacteria density is at first high (favouring the oxic state). The right column refers to when cyanobacteria density is at first low (favouring the anoxic state). Points (which sometimes are so close as to appear as thick lines) show the stable state at the end of each period of constant oxygen diffusivity. Thin lines are for visualisation only, and join the points. Grey arrows show the oxygen diffusivity at which the system flips, and the direction of the flip. The length of the grey arrows is without meaning.

@@

sim_length <- 1e6 ## currently a problem in how sim length is parsed by the data processing step
num_strains <- 9
plot_here <- all_stab %>%
  filter(simlength == sim_length) %>%
  ## remove data where there are non-standard stable states
## this subset is identified by inspection of stable state figures
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.70),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.0),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))

p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(sim_length,
                                                           " simulation length,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e6
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.0),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))

p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e6
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2

@@

wait_time <- 1e5
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

wait_time <- 1e4
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

```{r div-temp-meth4, echo = FALSE,  include=TRUE, fig.cap='(ref:div-temp-meth4)', fig.height=7, fig.width=6, fig.align="center"}
wait_time <- 1e3
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

```{r div-temp-meth5, echo = FALSE, eval = FALSE, include=TRUE, fig.cap='(ref:div-temp-meth5)', fig.height=7, fig.width=6, fig.align="center"}

## Owen removed these results from the report, because the tipping point finding algorithm was being misled
## by transient dynamics, and therefore was not finding the correct position of the anoxic to oxic tipping
## point

## (ref:div-temp-meth5) Effects of diversity with wait time of 1e2 hours. (All else as in Figure \@ref(fig:div-repl-meth1)).

wait_time <- 1e2
num_strains <- 9
plot_here <- all_stab %>%
  filter(waittime == wait_time) %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00))
p1 <- fig_div_vs_o2diff_1strain_7row(plot_here,
                                     which_strain = num_strains,
                                     figure_title = paste0(wait_time,
                                                           " wait time,\n",
                                                           num_strains,
                                                           " strains"))
p2 <- fig_resilience_vs_div(plot_here,
                            which_strain = num_strains,
                            figure_title = " ")
p1 + p2
```

@@

The results below show the stable states as a function of oxygen diffusivity for a system with an intermediate level of diversity in all functional groups, with either 1'000'000 hours at each oxygen diffusivity level (Figure \@ref(fig:ss-temp-meth3)), or 2'000'000 hours (Figure \@ref(fig:ss-temp-meth4)). This shows that 1'000'000 hours duration of the simulations are sufficient to reach stable state, except in regions of strain replacement, including at some state transitions. E.g. in the region of strain replacement in the sulfate-reducing bacteria recovery trajectory there is difference in strain abundances at 1'000'000 and 2'000'000 hours. However, the position of the tipping points (which were calculated from the total SB abundance) and the substrate concentrations did not differ between 1'000'000 and 2'000'000 hours.

@@

plot_here111 <- all_stab %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.0),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00),

         !(num_strains == 6 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 6 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 6 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 6 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 6 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 6 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 6 & var_treat == "CB-SB-PB" & stand_var > 1.00),

         !(num_strains == 3 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 3 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 3 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 3 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 3 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 3 & var_treat == "SB-PB" & stand_var > 1.0),
         !(num_strains == 3 & var_treat == "CB-SB-PB" & stand_var > 1.00),

         !(num_strains == 2 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 2 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 2 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 2 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 2 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 2 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == 2 & var_treat == "CB-SB-PB" & stand_var > 1.00))

fig_div_vs_o2diff_multistrain(plot_here111, which_strains,
                              which_wait_time, figure_title)
```

Increased CB diversity causes a slight decrease in the resilience of the anoxic state. The following two figures show the state responses to oxygen diffusivity with low Figure \@ref(fig:ss-CB-only1) and high Figure \@ref(fig:ss-CB-only2) levels of CB diversity. The difference in the position of the anoxic to oxic tipping point is small. It is, in fact the smallest it could be, as it is the smallest step size in oxygen diffusivity used in the simulations.

@@

(ref:pbdiv-only-3) Concentrations of three substrates (O: oxygen; SO: oxidised sulfur; SR: reduced sulfur) in a low PB diversity simulation (x-axis) versus substrate concentration in a moderate PB diversity simulation (y-axis). (The same levels of PB diversity as in Figures \@ref(fig:pbdiv-only-1) and \@ref(fig:pbdiv-only-2), respectively). The figures show only a small range of oxygen diffusivity on the trajectory of increasing oxygen diffusivity. The black line is the 1:1 line.

@@

dd %>%
  filter(State_variable %in% c("Oxygen", "Sulfide", "Oxidised sulfur"),
         direction == "up",
         a_O < as.numeric(aO_flip_sim2),
         a_O > -3.5) %>%
  ggplot(aes(x = log10(Quantity_sim1),
                 y = log10(Quantity_sim2),
                 col = a_O )) +
  geom_point(lwd = 3) +
  geom_path() +
  scale_color_gradient(low = "yellow", high = "red", na.value = NA,
                       breaks = labs,
                       labels = labs) +
  #coord_trans(x = "log10", y = "log10") +
  facet_wrap( ~ State_variable, scales = "free") +
  geom_abline(intercept = 0, slope = 1) +
  xlab(expression(atop(log[10](concentration)~(µM), Low~PB~diversity))) +
  ylab(expression(atop(log[10](concentration)~(µM),Moderate~PB~diversity))) +
  labs(col=expression(paste("log"[10],"(oxygen diffusivity) (h"^{-1}, ")"))) +
  theme_bw() +
  theme(legend.position="top")
```

@@

Owen, please add here 3 figures that match the text above:

```{r}
nine_strains <- readRDS(here("data",
                         microxanox_version_path,
                         "temporal_method", 
                         "event_definition_2",
                          "ss_data_9strains_waittime1e+06_event_definition_2.RDS"))
```

(1) medium variation in only PB showing coexistence of all groups at low ox diff,

nine_strain_one <- filter(nine_strains,
       CB_var_gmax_s == 0,
       SB_var_gmax_s == 0,
       abs(PB_var_gmax_s - 0.064819946) < 0.001)

parameter <- nine_strain_one$parameter[[1]]
parameter$strain_parameter <- nine_strain_one$pars[[1]]
#parameter$strain_parameter <- ss_result$pars[[1]]
total_initial_abundances <- 10^5
parameter <- set_temporal_ssfind_initial_state(parameter,
                                               total_initial_abundances,
                                               total_initial_abundances,
                                               total_initial_abundances)
parameter$log10a_series <- c(-7, -7)
wait_time <-  1e5
parameter$sim_duration <- wait_time * length(parameter$log10a_series)
parameter$sim_sample_interval <- 10
#parameter$sim_duration
test1 <- run_simulation(parameter)

(ref:PB-dyn1) Temporal dynamics of an ecosystem with a , the same one as in Figure \@ref(fig:max-tol3) simulated with constant $log_{10}$ oxygen diffusivity of -7.

```{r PB-dyn1, echo = FALSE,  include=TRUE, fig.cap='(ref:PB-dyn1)', fig.height=10, fig.width=8, fig.align="center"}
p <- plot_dynamics(test1)
p[[1]] <- p[[1]] + 
  xlab(NULL) + 
  ylab(expression(log[10](density)~(cells~L^{-1}))) +
  theme_bw()
p[[2]] <- p[[2]] + 
  xlab(NULL) + 
  ylab(expression(log[10](density)~(cells~L^{-1}))) +
  theme_bw()
p[[3]] <- p[[3]] + 
  xlab(NULL) + 
  ylab(expression(log[10](density)~(cells~L^{-1}))) +
  theme_bw()
p[[4]]$data <- p[[4]]$data %>% 
  dplyr::filter(species != "a")
p[[4]]$data$species <- p[[4]]$data$species %>%
  dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
p[[4]] <- p[[4]] + 
  xlab("time (h)") + 
  ylab(expression(log[10](concentration)~(µM))) + 
  labs(colour = "substrate") +
  theme_bw()

(2) high variation in only PB showing coexistence of all groups at low ox diff and CB never collapsing on the decreasing ox diff trajectory,

(3) high variation in both CB and PB.

@@

```{r samp-effect1, echo = FALSE, eval= FALSE, include=TRUE, fig.cap='(ref:samp-effect1)', fig.height=5, fig.width=6, fig.align="center"}

## no longer included in supplement.

## (ref:samp-effect1) A simple phenomenological model of the effect of number of strains in an ecosystem on the resilience of that ecosystem, when the resilience of a ecosystem is linearly proportional to the maximum trait value of the strains present in that ecosystem. Each data point is a different ecosystem, with the strains in each ecosystem drawn at random from a pool of 20 strains. Each of the 20 strains was assigned a trait value drawn at random from a uniform distribution with minimum 0 and maximum 1. The blue line is the fit of a generalised additive model and the grey ribbon shows the 1 standard error confidence range.

if(!("tidyverse" %in% .packages()))
  library(tidyverse)
set.seed(12345)
number_of_species <- 20
species_trait <- tibble(species_name = paste0("Species-", letters[1:20]),
                        trait_value = runif(number_of_species))

@@

We investigated if the diversity effects observed in our study were in line with the sampling effect, that is, driven entirely by the presence of the most tolerant strain or if other strains played a role as well. To this end, we ran additional simulations where functional groups contained only the strain with greatest tolerance (which also necessarily had the lowest maximum growth rate). That is, rather than increasing diversity we only increased tolerance of a single strain. Specifically, for each of the 20 levels of trait variation, we made simulations where each functional group had only one strain, and its trait values corresponded with those of the most tolerant strain for the given level of trait variation. We also manipulated in which of the functional groups tolerance increased. For example, in a simulation to be compared with variation in only the SB group (and no variation in the CB or PB groups), the SB group contained the strain with highest tolerance, whereas CB and PB each contained the strain with mean tolerance. We then compared the tipping point positions resulting from this tolerance manipulation with the tipping point positions resulting from the diversity manipulation presented in the main text. In addition, we compared the strain dynamics across the oxygen diffusivity gradient of the tolerance manipulation and the diversity manipulation. If the observed diversity effects are caused by the sampling effect, we expect there to be no difference between the tolerance manipulation and the diversity manipulation.

@@

## remove data where there are non-standard stable states
## this subset is identified by inspection of stable state figures
plot_here11 <- all_stab %>%
  filter(!(num_strains == 9 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "CB-SB" & stand_var > 1.00),
         !(num_strains == 9 & var_treat == "CB-PB" & stand_var > 0.65),
         !(num_strains == 9 & var_treat == "SB-PB" & stand_var > 1.0),
         !(num_strains == 9 & var_treat == "CB-SB-PB" & stand_var > 1.00),
         !(num_strains == -1 & var_treat == "CB" & stand_var > 1.00),
         !(num_strains == -1 & var_treat == "SB" & stand_var > 0.60),
         !(num_strains == -1 & var_treat == "PB" & stand_var > 1.00),
         !(num_strains == -1 & var_treat == "CB-SB" & stand_var > 0.60),
         !(num_strains == -1 & var_treat == "CB-PB" & stand_var > 1.00),
         !(num_strains == -1 & var_treat == "SB-PB" & stand_var > 1.00),
         !(num_strains == -1 & var_treat == "CB-SB-PB" & stand_var > 1.00))

plot_1 <- fig_div_vs_o2diff_maxtolcomp(plot_here11, which_strains,
                              which_wait_time, figure_title)
plot_1 +
  xlab("Standardised amount of trait variation used to create nine strains\nof which either all nine are retained, or only the most tolerant is retained.") +
  guides(col=guide_legend(title="Which strains are retained", title.position = "top"))
```

@@

```{r comp-comb, echo = FALSE, eval = FALSE, include=TRUE, fig.height=8, fig.width=8, fig.align="center", fig.cap='(ref:comp-comb)'}

## This code chunck makes a figure that we used but think does not need
## to be included in the supplement.
## (ref:comp-comb) Another method for comparing the effects of different diversity treatments. Some lines (treatment combinations) are over-plotted and therefore show identical effects of some combinations.

@@
